---
title: 'STAC67 Case Study: A predictive model that explains the median home value
  in Boston'
author: Group 21 / Eugenia Lee 1001715637 / Zhe Hao Xu 1002487962 / Daryl Chi Foong Yuen 1003043645 / Yuanzhen Shao 1003537274 / Bo Han Wang 1003287338 
date: "April 5, 2019"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

# Abstract
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; Housing values are valuable for home owners to determine the best suitable home and real estate agents to help clients. However, predicting the value of a house is a difficult task due to a plethora of factors that can cause the increase or decrease of its value. Some variables are better than others in predicting housing values. To determine which variables are the most significant, R is used for statistical analysis to find a relationship between median housing values in Boston and these influential factors.


```{r}
library(car)
library(MASS)
library("tidyverse")
library("SciViews")
```

```{r}
#load data
housing_proper = read_csv("housing.proper.csv")
# select random model building set and random validation set
smp_size <- floor(0.50 * nrow(housing_proper))
set.seed(123)
train_ind <- sample(seq_len(nrow(housing_proper)), size = smp_size)
housing_train = housing_proper[train_ind, ]
housing_validation = housing_proper[-train_ind,]
```


# Background and significance
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; The cost and quality of one's home are important factors that influence the quality of life in America. Housing represents the largest expense as well as the largest investment for most households (Burchell and Listoken, 1995). Thus, the demand for affordable housing is essential as it can be financially and emotionally taxing on the well-being of individuals and families (Bratt, 2002). People generally strive to live in an environment where it is safe to raise a family as well as provide quality schools (Jud, 1983). These are typically factors that are commonly known to affecting housing values. There are also other factors that could play significant roles in housing values such as the quality of air which is more difficult to quantify (Azmi et al., 2012). Ultimately, individuals and families strive to find housing that is not only within their budget but also allows for them to be in a comfortable environment whether it is the area they are located in or the house itself.

&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; However, housing is also a long-term investment (Burchell and Listoken, 1995). If an individual purchases a house, they want to know that their property has the ability to appreciate in value over time (Tighe, 2010). Having the knowledge of variables that can help explain increases and decreases in housing values is powerful information. It will can give insight for home buyers to be aware of the environment of the home.
      
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; There are an immense number of factors that come into play when determining the value of a house, some more significant than others. The data set given contains 506 observations on 13 variables that may affect the house value which in the data is the median value of owner occupied homes in thousands of dollars in the suburbs of Boston. The goal of this study is to determine the factors that best explains the variation of median housing values in the suburbs of Boston. We will be using various statistical techniques in R to compute the results.


# Exploratory data analysis
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;This data set contains 506 observations on 13 predictors that may affect the response variable which in this data is the median value of the owner occupied houses in thousands dollars (MEDV).

1. Per capita crime rate by town (x1)
This refers to the ratio of crimes for each person by town. The mean for this data is 3.613524 and mode is 0.01501.

2. Proportion of residential land zoned for lots over 25,000 sq.ft. (x2)
The mean and mode of this variable is 11.36364 and 0 respectively. A mode of 0 shows that most of the people in Boston do not own residence for over 25,000sq.ft.

3. Proportion of non-retail business acres per town (x3)
A non-retail business usually refers to business with no direct sales of goods to the public, such as factories. The region with high proportion of non-retail business acres per town relates to possibility of industry-noise and air pollution issues. The mean is 11.13678 and the mode is 18.1.

4. Charles River dummy variable (x4)
The Charles River dummy variable is a numeric variable set to be 1 if a home is near the river and 0 otherwise. The mean is 0.06916996 and mode is 0.

5. Nitric oxide concentration (parts per 10 million) (x5)
The nitric oxide concentration measured in parts per 10 million. The mean and mode is 0.5546951 and 0.538 respectively. If a person inhales or is exposed to Nitric oxide for long durations, it leads to severe health problems. 

6. Average number of rooms per dwelling (x6)
This variable refers to the average number of rooms in owners' units. The mean is 6.284634 and mode is 5.713.

7. Proportion of owner-occupied units built prior to 1940 (x7)
This variable refers to the age of the house. The age of a house is related to the quality of building's construction. The mean and mode is 68.5749 and 100 respectively. From the mode of this variable it shows most of the owners' house are built prior to 1940. 

8. Weighted distances to five Boston employment centers (x8)
From a recent report, it was shown that Boston was ranked as the third best city for job seekers in 2018 (Kimmel, 2018). Therefore, the demand for housing near employment center is essential. The mean is 3.795043 and the mode is 3.4952.

9. Index of accessibility to radial highways (x9)
The accessibility to radial highways relates to convenience of transportation. The mean and mode are 9.549407 and 24.

10. Full-value property-tax rate per 10,000 (x10)
This refers to the amount of tax that must be paid by the owner per $10,000. The mean is 408.2372 and mode is 666. This variable is affected by the value of property since the calculations are based on the housing values (Walsh, 2019).

11. Pupil-teacher ratio by town (x11)
This refers to the number of students for each teacher by town. A lower ratio implies more individual attention by teachers. The mean is 18.45553 and mode is 20.2.

12. 1000(B-0.63)^2 where B is the proportion of African Americans by town (x12)
The formula shows that when B increases the value of x12 decreases. The mean and mode are 356.674 and 396.9 respectively.

13. A numeric vector of percentage values of lower status population (x13)
The mean is 12.65306 and mode is 8.05. The data shows minor part of the population is of lower status. 


# Model
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; We split the 506 observations in half at random with 50% composing of the validation set and 50% composing of the model (building/training data set). We also transformed the response variable y (the median housing values) using the natural logarithm in order to make the data better follow a normal distribution. In order to determine our model, we performed backwards elimination by reducing one variable at a time, more specifically taking out the one with the smallest p-value while ensuring the \( R^2 \) value did not drop significantly. Using a significant level of 0.05 we were able to reduce the number of predictor variables from 13 to 10. We also took out variable x10 because our computed VIF tables showed that x10 has a high value of VIF and removing x10 results in a lower mean VIF value. In the end we reduced our model from 13 to 9 variables as shown in the summary below. Additionally to select our model, we applied Mallow's \(C_p\) criterion, comparing the \(C_p\) with the \(p\prime\) value. Our model generated a \(C_p\) value of 10 and its \(p\prime\) of 10 which indicates that residual sum of squares is an unbiased estimate of \( \sigma^2 \)(n-\(p\prime\)). To validate our model we compare MSPE and MSRes using our validaiton set. We obtained a MSRes of 19.39 and compared this value with the MSPE value of 18.63. This indicates a good model because an MSPE close to the MSRes based on the regression fit to the training data set indicates good predictive ability of the model. 

# Summary of Model
```{r}
model <- lm(log(y) ~ x1 + x4 + x5 +x6 +x8 +x9 +x11 +x12+ x13, data = housing_train)
summary(model)
```


# Diagnostics

## Residuals vs Fitted
```{r}
final_model <- lm(log(y) ~ x1 + x4 + x5 +x6 +x8 +x9 +x11 +x12+ x13, data = housing_proper)
```
```{r}
## diagnostic plots
plot(final_model,which = 1, ask = FALSE)
```
The residual plot shows a random pattern which indicates that our selected model has proper functional form.

## Normal Q-Q Plot
```{r}
## diagnostic plots
plot(final_model,which = 2, ask = FALSE)
```

The normal Q-Q plot shows that the errors are not relatively normal and the normality assumption does not hold based on the plot. Though, with a larger dataset of 506 we can assume that the normality assumption is not needed based on the Central Limit Theorem which states that the distribution of the errors will be approximated to normal.

## Multicollinearity
```{r}
vif_table <- vif(final_model)
vif_table
VIFbar <- mean(vif(final_model))
VIFbar #VIF mean
```
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; We calculated the variance inflation factor (VIF) to investigate multicollinearity between variables in our final model. We found that there are no indiciations of serious multicollinearity because all the obtained VIF values are below 10. As such, none of our variables in our final model are colinear, meaning our model is valid by the effects of the violation multicollinearity.


## Outlying Y values
```{r}
outlierTest(final_model)
t <- rstudent(final_model)
alpha <- 0.05
n <- 506
p_prime <- length(coef(final_model))
t_crit <- qt(1 - alpha / (2 * n), n - p_prime - 1)
t_crit
```

Analysis of outlying Y obervations showed that observation 413 and 372 are outliers. This is because our obervations have a larger studentized residual than the \(t_{critical}\) value of 3.92

## Outlying X Observations
```{r}
Pii <- hatvalues(final_model)
n <- 506
#round(Pii, 2)
guideline1 <- which(Pii > 2*p_prime/n)
guideline2 <- which(Pii > 0.5)

guideline1
guideline2
```
We tested for leverage points using the two guidelines \(P_{ii}\) > 2*\(p_{prime}\)/n and \(P_{ii}\) > 0.5. We found that using guideline 1 we found 57 observations and guideline 2 we found no observations. We used guideline 2 in the end because it is less sensitive than guideline. As such, we conclude that there are no significant outlying x observations.


## Influence
```{r}

## Influence 
influencePlot(final_model, main="Influence Plot", sub="Circle size is proportial to Cook's Distance" )

```
Our influence plot shows 5 obervations with high Cook's distances (372, 381, 406, 413 and 419). We also calculated the 10th, 20th, and 50th percentile of F(10, 496) resulting in 0.485, 0.617 and 0.935 respectively. We see that the Cook's distances for our obervations are less than the 10th, 20th and 50th percentile meaning that all these observations have little inflence on the fitted values. Due to little influence, we decided to keep these obervations in our dataset when making our model. To further analyze infleuntial points, we looked at \(DFBETAS_{k(i)}\) and found that the obervations were not influential since they were less than 1. 



# Discussion/Conclusions
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; The goal of this study was to derive a model based on the dataset that is able to predict mean housing values in the suburbs of Boston. Using the natural logarithm of median housing values, we found that the variables that best explain the value are crime rate, being close to Charles River, nitric oxide concentration, average number of rooms per dwelling, distance from employment centres, accessibility to highways, the ratio of pupil to teachers and the proportion of African Americans. Real estate companies can benefit from this research in using this model to predict housing values. Home owners can also benefit from this model by seeing which variables they should be looking at when making investments in a home. The limitations of this study is that the derived model can only be applied to homes in Boston, Massachusetts and more specifically in the suburbs. Possible areas of future study could include non-suburban areas to see the effects of these same variables and if they still apply in that context. 

# References

Azmi, A. S., Azhar, R. F., & Nawawi, A. H. (2012). The Relationship Between Air Quality and Property Price. Procedia - Social and Behavioral Sciences,50, 839-854. doi:10.1016/j.sbspro.2012.08.086

Burchell, R. W., & Listokin, D. (1995). Influences on United States housing policy. Housing Policy Debate,6(3), 559-617. doi:10.1080/10511482.1995.9521197

Jud, G. D. (1983). Schools and Housing Values: Reply. Land Economics,59(1), 135. doi:10.2307/3145887

Kimmel, J. (2018). Boston one of the top cities in the United States for job seekers. report says. (2018, April 26). Retrieved from https://dailyfreepress.com/blog/2018/04/26/boston-one-of-the-top-cities-in-the-united-states-for-job-seekers-report-says/

Tighe, J. R. (2010). Public Opinion and Affordable Housing: A Review of the Literature. Journal of Planning Literature,25(1), 3-17. doi:10.1177/0885412210379974

Walsh, M. J. (2019, January 01). How we tax your property. Retrieved April 05, 2019, from https://www.boston.gov/departments/assessing/how-we-tax-your-property





 


